---
title: "wikitablr Vignette"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{my-vignette}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```

```{r setup, include=FALSE}
library(wikitablr)
library(dplyr)
library(stringr)
```

There are two main families of functions in `wikitablr`:

1. Those that read in a table and
2. Those that clean a table

There is much conjunction within these categories, and there are functions that do both. These are `read_all_tables()`, which reads in and cleans every table at the specified Wikipedia page and returns them in a list, and `read_wiki_table()`, which reads in and cleans a specified table (designated by the table number).

```{r}
#for this vignette I will be using the wikipedia page of Bob's Burgers episodes
#this is a table of all episodes in season 10
read_wiki_table("https://en.wikipedia.org/wiki/List_of_Bob%27s_Burgers_episodes", table_number = 11) %>% head(3)
```


However, users do have the option to read and clean their table seperately. In this vignette, I detail all of my cleaning functions, and show how they are used in `read_wiki_table()`.

A table can be read in without cleaning using `read_wiki_raw()`.

```{r}
episodes <- read_wiki_raw("https://en.wikipedia.org/wiki/List_of_Bob%27s_Burgers_episodes", 11)
head(episodes, 3)
```

#### Clean column names

`clean_wiki_names()` cleans the column names of a dataframe; it removes footnotes and special characters, and converts letter cases. `clean_wiki_names()` utilizes the functionality of `janitor::clean_names()` and can convert to any case (default is snake_case).

Here, we look at the column names of the uncleaned data:

```{r}
colnames(episodes)
```

And compare it to the cleaned column names.

```{r}
episodes <- clean_wiki_names(episodes)
colnames(episodes)
```


#### Add NAs to data
`add_na()` takes blank cells, solitary special characters, and any designated string, and converts them to `r NA`. This can be utilized on the table for Bob's Burgers' current season, which has some "TBA" and "TBD" values for unreleased episodes. Using `add_na()` we can convert these to `r NA`:

```{r}
episodes <- add_na(episodes, to_na = c("TBA", "TBD"))

tail(episodes, 3)
```

#### Remove footnotes
`remove_footnotes()` does just that! Wikipedia's footnotes are contained within brackets ([ ]), so this function removes anything contained in brackets (as well as the brackets themselves) from the data.

```{r}
episodes <- remove_footnotes(episodes)

head(episodes, 3)
```

#### Convert data types
`wikitablr` also does some initial parsing of data types. `convert_type()` can detect if a column is likely Date or numeric, and re-codes it as such:

```{r}
str(convert_types(episodes))
```

As can be seen, this function isn't perfect, and the original_air_date column was parsed into character, rather than Date. A situation like this calls for some manual data-cleaning:

```{r}
#using stringr to remove the duplicate date in parentheses
episodes$original_air_date <- stringr::str_remove(episodes$original_air_date, "\\(.*\\)") %>% stringr::str_trim()

str(convert_types(episodes))
```

Now `convert_type()` correctly recodes original_air_date as Date!

#### Remove unnecessary rows

`wikitablr`'s final function, `clean_rows()` removes unnecessary rows: those that are duplicates of the header, and those that have the same value for every column (seen when the Wikipedia table has a banner in the middle).

To illustrate the first purpose, I use a table of US presidents by age.
```{r}
presidents <- read_wiki_raw("https://en.wikipedia.org/wiki/List_of_presidents_of_the_United_States_by_age", 2)

tail(presidents, 3)
```

As can be seen, the header is duplicated at the bottom row. This is a quirk not uncommmon in Wikipedia tables; however, it is quickly and easily solved with `clean_rows()`.
```{r}
clean_rows(presidents) %>% tail(3)
```

Wikipedia's other quirk, banners in the middle of tables, is seen in this list of Marvel Cinematic Universe films:

```{r}
marvel <- read_wiki_raw("https://en.wikipedia.org/wiki/List_of_Marvel_Cinematic_Universe_films")

head(marvel, 3)
```

Here, the row of "Phase One" could interfere with cleaning and analysis, so it is best to remove it with `clean_rows()`

```{r}
clean_rows(marvel) %>% head(3)
```

### Clean all at once

`read_wiki_table()` utilizes all of these cleaning functions internally, so you don't have to!

```{r}
read_wiki_table("https://en.wikipedia.org/wiki/List_of_Bob%27s_Burgers_episodes", 11) %>% head(3)
```

Because it calls on the above functions, `read_wiki_table()` takes the same arguments as well.

```{r}
read_wiki_table("https://en.wikipedia.org/wiki/List_of_Bob%27s_Burgers_episodes", 11, to_na = "Mario D'Anna", case = "all_caps") %>% head(3)
```

### Less common scenarios

Although `wikitablr` cannot account for all the inconsistencies found in Wikipedia tables, here are a few more that you might come across:

* both `read_wiki_table()` and `read_wiki_raw()` have the `replace_linebreak` parameter. This parameter determines what is input in place of a linebreak is cell. The default is `", "`, but users can change to any other character string desired.
```{r}
read_wiki_table("https://en.wikipedia.org/wiki/List_of_songs_recorded_by_the_Beatles", 2, replace_linebreak = "\n") %>% head(3)
```

* `add_na()` converts solitary special characters to NA by default, but this can be overridden.
```{r}
read_wiki_raw("https://en.wikipedia.org/wiki/List_of_World_Series_champions", 2) %>% add_na() %>% tail(3)

read_wiki_raw("https://en.wikipedia.org/wiki/List_of_World_Series_champions", 2) %>% add_na(special_to_na = FALSE) %>% tail(3)
```

